### 2020-08-06 paper.tex
Rewrite Interivew/War Story Quotes

All of the interviewees referred to RP with an overall positive experience. We heard statements like ``RP is a good way for composing multiple data sources'' and ``having a statically typed language combined with RP already guarantees some kind of basic correctness of my program'' repeatedly throughout all sessions. Hence a significant strength of RP seems to be the ability to describe complex dataflow constructs using a specialized DSL. The learning curve can be steep for a novice engineer though: ``Being already challenged with new concepts of Angular and ngrx, I noticed RxJS concepts and operators to be hard to grasp in addition'' an interview partner giving lectures in frontend web application development stated further.

It was interesting to hear that, especially in the area of developing web applications using Angular, our partners seemed not to have to work with pure RxJS code often. E.g., when using ngrx for state management, ``The framework takes care that you do not have to interact with Observables on your own, but leverages greatly on their potential behind the scenes''. As soon as our interviewees had to extend built-in functionalities with their own features, e.g., a new effect\footnote{\url{https://ngrx.io/guide/effects}}, they all valued the possibility to interact with underlying Observables highly.

When asked explicitly about what they dislike the most about RP, all interview partners, with no exception, emphasized the debugging process of an RP application as a big nuisance. The fact that our interviewees remembered the search for a bug as something negative did not surprise us, hence a bug is usually something negative afflicted. What did alarm us though was the fact that statements like ``In 99\% of all cases, I add console.log statements manually and run the program over and over again, trying to understand what is happening'' were prevalent and gave insight in \emph{why} our interview partners dislike RP debugging in particular. Multiple interviewees mentioned the Redux DevTools\footnote{\url{https://chrome.google.com/webstore/detail/redux-devtools}} as helpful when debugging Angular/ngrx applications nonetheless. Further, one single partner mentioned marble diagrams as valuable in order to understand how an RxJS Observable works, whether during development or when debugging code. Marble diagrams are a way to visualize and describe the value-emitting behavior of an Observable over time. They are accepted throughout the ReactiveX community for documentation\cite{marblediagrams} and testing reasons\cite{marbletesting}.

@@

The reports provided us with diverse perspectives on various areas of software development using RxJS. One of the reporters focused on how they built code with improved testability because of recent changes in RxJS: ``[\dots] in the beginning, it was very hard to write async tests and I did not like it. I really disliked that once you started to test async code you were forced to pass a TestScheduler''. Allowing to pass the scheduler explicitly as parameter forced them to introduce code, which was only necessary for testing reasons they stated further. With its current major version, RxJS 6 improved heavily on the  \mintinline{TypeScript}{TestScheduler}. The runtime environment itself can be augmented with the scheduler now, which results in cleaner code.

Even though the share of non-productive, testing-related code necessary to build mature RxJS-based applications was mentioned to have decreased today, multiple of our reporters commented on the common practice of manually modifying production code during debugging sessions, hence confirming earlier statements from our informal interview partners. One engineer described a specific scenario where they suspected a problem within a complex Observable composition: Having multiple asynchronous, remote data sources, they used Observables to model the dependencies between them and implemented computations on their results as operators. On top, each data source could re-emit updated versions of previously requested information at any time. Though tested thoroughly, after a week in production, the first of many bugs were reported: ``Figures were changing multiple times (where I expected them to change once), other numbers were not updated at all,'' they told us. The browsers debugger tools and its breakpoints did not help much since the operators were executed several multiple times and some parts of the stream were impossible to grasp, even with conditional breakpoints. ``I started to inspect the flow [\dots] with console.logs and later also using tags from rxjs-spy which exposed more detailed life cycle information'', hence manually augmenting the code with trace logs. After the time-consuming log analysis, they finally were able to resolve the bug. The log statements added were removed in the aftermath, though the rxjs-spy related code was left in the Observable stream in case they might be needed again in the future. (We will elaborate on the functionality of rxjs-spy in Section~\ref{sec:discussion} more in-depth.)

Another engineer disclosed similar practices using external tools in a related war story: ``It took quite some time to get it right and one of the most invaluable tools proved to be Stackblitz which gave us ability to quickly create smaller working examples and iterate on them.'' Stackblitz\footnote{\url{https://stackblitz.com/}} is a full JavaScript development environment available online. They used this sandboxed setting to run, debug, and iterate on selected pieces of a larger Observable stream. Even though the final result had to be integrated back into the actual application, the extra effort was worth the result the engineer concluded in their report.

A final story described yet another way of utilizing an external tool: Rather than using a dedicated sandbox to develop pieces of a more complex system, this engineer used Rx Visualizer, an online visualization utilities to generate previously presented marble diagrams from real code. Like in the report before, it was necessary to extract parts from the codebase of the actual application. But once done, the visualizations helped a lot in order to understand when values were emitted, when subscriptions changed and when Observables completed: ``Marble diagrams were a huge help to dig into detailed runtime and life cycle behaviors of the Observables.''.

>>>>

All of the interviewees referred to RP with an overall positive experience. We heard statements similar to ``RP is a good way for composing multiple data sources'' and ``having a statically typed language combined with RP already guarantees some kind of basic correctness of my program'' repeatedly throughout all sessions. Hence a significant strength of RP seems to be the ability to describe complex dataflow constructs using a specialized DSL. The learning curve can be steep for a novice engineer though: ``Being challenged with new abstractions [of Angular and ngrx] already, I experienced RxJS concepts and operators to be hard to convey'' an interview partner giving lectures in frontend web application development stated further.

It was interesting to hear that, especially in the area of developing web applications using Angular, our partners seemed not to have to work with pure RxJS code often. E.g., when using ngrx for state management, ``The framework hides Observables from its main API surface carefully so you do not have to interact with them directly''. As soon as our interviewees had to extend built-in functionalities with their own features, e.g., a new effect\footnote{\url{https://ngrx.io/guide/effects}}, they all valued the possibility to interact with underlying Observables highly.

When asked explicitly about what they dislike the most about RP, all interview partners, with no exception, emphasized the debugging process of an RP application as unsatisfactory. The fact that our interviewees remembered the search for a bug as something negative did not surprise, hence a bug is commonly something negative afflicted. It was remarkable though that statements like ``In 99\% of all cases, I add console.log statements manually and run the program over and over again, trying to understand what is happening'' were prevalent and showed \emph{why} our interview partners dislike RP debugging in particular. Multiple interviewees mentioned the Redux DevTools\footnote{\url{https://chrome.google.com/webstore/detail/redux-devtools}} as helpful when debugging Angular/ngrx applications nonetheless. Further, one single partner mentioned marble diagrams as valuable in order to understand how an RxJS Observable works, whether during development or when debugging code. Marble diagrams are a way to visualize and describe the value-emitting behavior of an Observable over time. They are accepted throughout the ReactiveX community for documentation\cite{marblediagrams} and testing reasons\cite{marbletesting}.

@@

The reports provided us with diverse perspectives on various areas of software development using RxJS. One of the reporters focused on how they built code with improved testability because of recent changes in RxJS: ``[\dots] in the beginning, it was very hard to write asynchronous tests [\dots]. I really disliked that, once you started to test asynchronous code, you were forced to pass a TestScheduler''. Allowing to pass the scheduler explicitly as parameter forced them to introduce code, which was only necessary for testing reasons they stated further. With its current major version, RxJS 6 improved heavily on the  \mintinline{TypeScript}{TestScheduler}. The runtime environment itself can be augmented with the scheduler now, which results in cleaner code.

Even though the share of non-productive, testing-related code necessary to build mature RxJS-based applications was mentioned to have decreased today, multiple of our reporters commented on the common practice of manually modifying production code during debugging sessions, hence confirming earlier statements from our informal interview partners. One engineer described a specific scenario where they suspected a problem within a complex Observable composition: Having multiple asynchronous, remote data sources, they used Observables to model the dependencies between them and implemented computations on their results as operators. On top, each data source could re-emit updated versions of previously requested information at any time. After a week in production, though tested thoroughly, the first of many bugs got reported: ``Displayed information kept changing/flickering (where we expected it to change exactly once and stay), other data got not updated at all when it was supposed to,'' they told us. The browsers debugger tools and its breakpoints did not help much since the operators were executed several multiple times and some parts of the stream were impossible to grasp, even with conditional breakpoints. ``I started to inspect the flow [\dots] with console.logs and later also using tags from rxjs-spy which exposed more detailed life cycle information'', hence manually augmenting the code with trace logs. After the time-consuming log analysis, they finally were able to resolve all bugs. The log statements added were removed in the aftermath, though the rxjs-spy related code was left in the Observable stream in case they might be needed again in the future. (We will elaborate on the functionality of rxjs-spy in Section~\ref{sec:discussion} more in-depth.)

Another engineer disclosed similar practices using external tools in a related war story: ``It took quite some time to get it right and one of the most invaluable tools proved to be Stackblitz\footnote{Stackblitz is a full JavaScript development environment available online \url{https://stackblitz.com/}} which gave us ability to quickly create smaller working examples and iterate on them.'' This sandboxed setting allowed P999 to run, debug, and iterate on selected pieces of a larger Observable stream. Even though the final result had to be integrated back into the actual application, the extra effort was worth the result the engineer concluded in their report.

A final story described yet another way of utilizing an external tool: Rather than using a dedicated sandbox to develop pieces of a more complex system, this engineer used Rx Visualizer, an online visualization utilities to generate previously presented marble diagrams from real code. Like in the report before, it was necessary to extract parts from the codebase of the actual application. But once done, the visualizations helped a lot in order to understand when values were emitted, when subscriptions changed and when Observables completed: ``Marble diagrams were a huge help in order to understand detailed runtime and life cycle behaviors of the Observable.''
